---
title: Web Traffic Forecasting
subtitle: Time Series Analysis
output:
  pdf_document:
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Team Members (Group 11)

* Jackson Smith
  + Email:
  + Major:
  + Responsibility:
* Jingcheng Xia
  + Email:
  + Major:
  + Responsibility:
* Max Kutschinski
  + Email: mwk556@tamu.edu
  + Major: Statistics (Distance)
  + Responsibility:
* Samuel Burge
  + Email: samuelburge@tamu.edu
  + Major: Statistics (Distance)
  + Responsibility:
* Shree Karimkolikuzhiyil
  + Email:
  + Major:
  + Responsibility:


\newpage

## Introduction

<!-- Introduce Your Data: Explain the context and background story of your data -->

Why web traffic forecasting is important:

- optimize web server allocation
- helps enterprises plan their IT infrastructure strategy
- facilitates seamless end user experience and satisfied customers

About the dataset: 

- contains 5 years of daily time series data of user visits to an academic website
- Recorded variables: 
   - Number of page loads
   - Number of unique visitors (unique visitors are identified through IP address within a 6 hour window)
   - Number of returning visitors (identified by cookies if accepted)
   - Number of first-time visitors
- Hence number of unique visitors is the number of returning visitors + number of first-time visitors.
- The data was collected through a traffic monitoring service known as StatCounter.




<!-- Plot the Data and summarize your preliminary findings: trend, cycle, volatility, etc. --> 

```{r include=F, echo=F}
library(readr)
library(astsa)
library(ggplot2)

webData = read_csv("WebTraffic.csv",
                   skip = 1,
                   col_names = c("Day",
                                 "DayOfWeek",
                                 "Date",
                                 "PageLoads",
                                 "UniqueVisits",
                                 "FirstTimeVisits",
                                 "ReturningVisits"),
                   col_types = cols("Day" = "f",
                                    "Date" = col_date("%m/%d/%Y")))
```

```{r include=F, echo=F}
colnames(webData)
anyNA(webData)
dim(webData)
str(webData)
summary(webData)
```

```{r echo=F, fig.width = 18, fig.height=7}
theme_set(theme_bw())
ggplot(webData,aes(Date, UniqueVisits))+
  geom_line()+
  scale_x_date(date_labels = "%Y", date_breaks = "1 year")+
  labs(title = "Unique Page Visits Over Time")+ 
  ylab("Unique Visits")+
  theme(text = element_text(size = 30)) 

```

- strong time of year effect






