---
title: \vspace{-2.0cm}Web Traffic Forecasting
subtitle: Time Series Analysis
author:
- \small Shree Karimkolikuzhiyil; Statistics, M.S. ([shreejesh\@tamu.edu](mailto:shreejesh@tamu.edu){.email})
- \small Jingcheng Xia; Computer Science, B.S.  ([sixtyfour64\@tamu.edu](mailto:sixtyfour64@tamu.edu){.email})
- \small Jackson Smith; Statistics, M.S. ([jackson.t.smith\@tamu.edu](mailto:jackson.t.smith@tamu.edu){.email})
- \small Samuel Burge; Statistics, M.S. ([samuelburge\@tamu.edu](mailto:samuelburge@tamu.edu){.email})
- \small Max Kutschinski; Statistics, M.S. ([mwk556\@tamu.edu](mailto:mwk556@tamu.edu){.email})



output:
  pdf_document
fontsize: 11pt
---
\vspace{-1.5cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# package to use the Latex font in plots
library(showtext)
font_add(family = "cm", regular = "computermodern.ttf")
showtext_auto()
```
---
## Introduction and Motivation

<!-- Introduce Your Data: Explain the context and background story of your data -->
The objective of this analysis is to forecast daily unique visitors to an academic website containing lecture notes and supplemental material related to statistics. Predicting website traffic allows IT departments to manage project throughput and prioritize maintenance and enhancements to website functionality and effectively allocate web server resources. Web traffic is also a key indicator of customer growth and expansion, as well as sustaining recurring customers and ingrained growth. The details provided by web traffic throughput reports contain many metrics, including page loads, returning visitors, and unique visits, each of which convey a different picture and set of information for an organization. As well, having a picture of expected throughput and confirming (or denying) expectations with reality allows a business to understand unexpected growth and/or unexpected decay in business development.

The data contains five years of daily time series data of user visits. There are four features in the data set, which are the daily counts for the number of page loads, first-time visitors, returning visitors, and unique visitors.^[Returning visitors are identified through allowed cookies on a user's device, and the total returning and first-time visitors is, by definition, the number of unique visitors. Users are identified through the IP address within a 6 hour window.] An initial plot of the data shows strong seasonality and volatility, but doesn't appear to have any discernible trend or cyclical behavior. An explanation for this could be due to the nature of the website. Students would likely be the largest share of users for a website of this nature, and the seasonality seems associated with the academic calender typically seen at academic institutions.  
\  

<!-- Plot the Data and summarize your preliminary findings: trend, cycle, volatility, etc. -->
```{r include=F, echo=F}
library(readr)
library(astsa)
library(ggplot2)

webData = read_csv("WebTraffic.csv",
                   skip = 1,
                   col_names = c("Day",
                                 "DayOfWeek",
                                 "Date",
                                 "PageLoads",
                                 "UniqueVisits",
                                 "FirstTimeVisits",
                                 "ReturningVisits"),
                   col_types = cols("Day" = "f",
                                    "Date" = col_date("%m/%d/%Y")))
```

```{r include=F, echo=F}
colnames(webData)
anyNA(webData)
dim(webData)
str(webData)
summary(webData)
```  


```{r echo=F, fig.width = 17, fig.height=6}

# Set the theme for the chart
theme_set(theme_bw() + theme(text = element_text(family = 'cm')))

# Plot the daily count of unique page visits
ggplot(webData,aes(Date, UniqueVisits)) +
  
  geom_line(color = '#437cb5') +
  
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  
  labs(title = "Unique Page Visits Over Time") + 
  
  ylab("Unique Visits") +
  
  theme(text = element_text(size = 30)) 

```



